---
title: "Farm-level balance and strip-prescription map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Farm-level balance and strip-prescription map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
```

This vignette walks through the end-to-end workflow that NFert v0.13
supports:

  1. Load a farm layout (GeoJSON) with per-plot agronomic inputs.
  2. Run the DPI-2026 N balance on every plot.
  3. Build a machine-width strip prescription map for one plot.
  4. Export the prescription in the formats accepted by the on-board
     monitor of the target tractor (Shapefile, ISOXML TASKDATA.XML,
     John Deere / Trimble flavours, etc.).

## 1. Load the demo farm

The package ships with a fictitious 35.5 ha farm near Piacenza with 8
plots carrying all the columns required by `N_balance()`:

```{r eval = requireNamespace("sf", quietly = TRUE)}
library(NFert)
library(sf)
ex   <- system.file("extdata/example_farm.geojson", package = "NFert")
farm <- sf::st_read(ex, quiet = TRUE)
farm[, c("plot_id", "plot_name", "crop", "area_ha")]
```

## 2. Per-plot N balance

`farm_balance()` applies `N_balance()` to every feature and appends
`N_target`, `MAS_cap`, `MAS_ok` and `N_total_kg` to the layer. Italian
legacy strings are accepted in input and translated on the fly.

```{r eval = requireNamespace("sf", quietly = TRUE)}
farm_out <- farm_balance(farm)
farm_out[, c("plot_id", "crop", "N_target", "MAS_cap",
             "MAS_ok", "N_total_kg")]
```

## 3. Strip prescription along an A-B line

`build_strip_prescription()` tiles one plot with parallel strips of a
user-supplied working width. If no A-B line is given, the longest side
of the field is used as driving direction. Dose variability can be
driven by a uniform target, a VI calibration curve, NNI zones, or
quantile classes.

```{r eval = requireNamespace("sf", quietly = TRUE), fig.height = 5}
field <- farm[1, ]          # Campo Grande, 5.2 ha silage maize

# Option A: uniform 226 kg N/ha (from the balance), 24 m spreader
rx_uni <- build_strip_prescription(
  field         = field,
  machine_width = 24,
  variability   = "uniform",
  n_target      = 226)

plot(sf::st_geometry(rx_uni), col = "#D5E4F0",
     border = "#1F4E79", main = "Uniform strips, 24 m")
```

Swapping in a VI calibration curve makes the dose decrease with
vigour; the mean across the field is preserved (mass-balance
constraint) by default.

```{r eval = FALSE}
# Option B: VI calibration (requires an NDVI raster)
rx_cal <- build_strip_prescription(
  field         = field,
  machine_width = 24,
  variability   = "calibration",
  vi_raster     = my_ndvi_raster,
  n_target      = 226,
  min_dose      = 80,  max_dose = 260,
  vi_low        = 0.35, vi_high = 0.80)
```

A 2-D **grid** (variable dose cell-by-cell) is produced by passing
`cell_length > 0`; an explicit A-B direction is set with
`angle_deg` (degrees, 0 = east, 90 = north) or by supplying an
`sf` LINESTRING in `ab_line`:

```{r eval = FALSE}
# Option C: 24 m working width x 50 m along-strip cells, AB = 30 deg
rx_grid <- build_strip_prescription(
  field         = field,
  machine_width = 24,
  cell_length   = 50,
  angle_deg     = 30,
  variability   = "classes",
  vi_raster     = my_ndvi_raster,
  n_target      = 226,
  min_dose      = 80, max_dose = 260,
  n_classes     = 5)

# Option D: NNI-driven zones
rx_nni <- build_strip_prescription(
  field         = field,
  machine_width = 24,
  cell_length   = 0,           # continuous strips
  variability   = "nni",
  nni_raster    = my_nni_map,
  n_target      = 226,
  min_dose      = 80, max_dose = 260,
  thr_lo        = 0.90, thr_hi = 1.10)
```

## 4. Export to tractor-ready formats

`export_prescription()` writes a single file; `export_prescription_all()`
bundles several formats in one output folder.

```{r eval = FALSE}
export_prescription(rx_uni, "rx.shp",       format = "shp")
export_prescription(rx_uni, "JD/rx.shp",    format = "johndeere")
export_prescription(rx_uni, "TASKDATA",     format = "isoxml",
  isoxml_opts = list(task_name = "N top-dress",
                      product   = "Urea 46 pct",
                      unit      = "kg/ha"))

export_prescription_all(rx_uni, "rx_bundle", "campo_grande",
  formats = c("shp", "isoxml", "johndeere"))
```

The Shiny interface (`NFert::run_app()`) exposes the same flow with
interactive forms and a leaflet preview.
