---
title: "Nitrogen Balance"
author: "Michele Croci, Stefano Amaducci"
contact: "michele.croci@unicatt.it"
institute: "Department of Sustainable Plant Productions (DI.PRO.VE.S), Catholic University of the Sacred Heart"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Nitrogen Balance Calculation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

## Overview

The `N_balance()` function is a comprehensive tool for calculating the
nitrogen balance in agricultural systems. It takes into account various
soil characteristics, crop specifications, and environmental factors to
provide a detailed analysis of nitrogen dynamics in crop production.

## Course Objectives

1.  **Understanding and Applying Nitrogen Balance** in Herbaceous Crops
2.  **Developing Skills** in Water Balance Calculation
3.  **Introducing Precision Agriculture** Concepts and Its Impact on
    Agricultural Practices

## Soil Sampling

### Sampling Period

The optimal soil sampling period should consider: - Soil moisture status
(neither too dry nor too wet) - Sufficient time after tillage and
fertilization operations - For herbaceous crops, ideal timing is: - Days
following harvest - At least two months after the last fertilizer
application

### Sampling Methodology

To obtain a representative sample:

-   Proceed in a zigzag pattern across the field
-   Select 15-20 sampling points depending on field extension
-   Remove the first 5 cm to eliminate surface vegetation and debris
-   Collect samples to a depth of 30 cm (0-30 cm)
-   Thoroughly mix the collected soil
-   Remove stones and coarse organic matter

```{r}
library(NFert); library(raster)
```


## Function Parameters

The function accepts the following key parameters:

-   `expected_yield_tons_ha`: Expected crop yield (tons per hectare)
-   `crop`: Name of the crop being cultivated
-   `ccp`: Crop calendar period
-   `sand`: Percentage of sand in the soil
-   `clay`: Percentage of clay in the soil
-   `Ntot`: Total nitrogen content in the soil (%)
-   `SOM`: Soil organic matter content (%)
-   `CN`: Carbon to nitrogen ratio of the soil

### Minimum Required Physico-Chemical Analyses

-   Granulometry (soil texture)
-   Water pH
-   Organic matter content (%)
-   Total and active limestone
-   Total nitrogen (g/kg)
-   Exchangeable potassium
-   Assimilable phosphorus

## Nitrogen Fertilization Calculation

The nitrogen fertilization is calculated using the following analytical
balance:

$$
\text{Nitrogen Fertilization (N)} = A - B + C + D - E - F - G
$$

Where: - **A**: Crop Requirements - **B**: Soil Fertility
Contributions - **C**: Leaching Losses - **D**: Immobilization and
Dispersion Losses - **E**: Nitrogen from Previous Crop Residues - **F**:
Nitrogen from Organic Fertilizations in Previous Years - **G**: Natural
Contributions

## Example Usage
```{r}
head(NFert::uptake_table, 15)

```

```{r}
# Maize crop in a typical agricultural setting
maize_nitrogen_balance <- N_balance(
  expected_yield_tons_ha = 15, 
  crop = "Shredded corn class 700",
  ccp = "Spring-summer crop 100–130 days", 
  sand = 50, 
  clay = 35,
  Ntot = 1.2, 
  SOM = 1.2, 
  CN = 9.5, 
  oxygen_availability = "Normal",
  winter_rain = 160, 
  start_spring_rain = 40,
  prev_crop = "Winter cereals straw removal", 
  source = "Cattle slurry",
  fertorg_frequency = "every year", 
  location = "Plain adjacent to urbanized areas",
  forg_quantity = 100
)

# View the nitrogen balance components
print(maize_nitrogen_balance)
```

``` r
```

## Key Nitrogen Balance Components

1.  **A**: Total crop nitrogen demand
2.  **B**: Soil nitrogen supply
3.  **b1**: Readily available mineral nitrogen
4.  **b2**: Mineralizable nitrogen from soil organic matter
5.  **C1**: Winter nitrogen leaching loss
6.  **C2**: Spring nitrogen leaching loss
7.  **D**: Nitrogen immobilization loss
8.  **E**: Nitrogen from previous crop residues
9.  **Forg**: Nitrogen from organic fertilizer
10. **G**: Natural nitrogen contribution

## Interpretation Guidelines

-   Positive nitrogen balance indicates potential nitrogen surplus
-   Negative balance suggests the need for additional fertilization
-   Balance helps optimize fertilizer application
-   Minimizes environmental impact
-   Supports sustainable agricultural practices

## Practical Considerations

-   Adjust parameters based on local soil conditions
-   Consider crop-specific nitrogen requirements
-   Monitor soil health and organic matter content
-   Use in conjunction with comprehensive soil management practices

## Limitations and Recommendations

-   Requires accurate input of soil and crop parameters
-   Represents a simplified model of complex nitrogen cycles
-   Local validation is strongly recommended
-   Reassess soil analysis every 5 years

## Natural Nitrogen Contributions

Nitrogen reaches the soil through: - Atmospheric precipitation -
Symbiotic nitrogen fixation by leguminous crops

Average annual nitrogen deposition: - Near urbanized areas: 20 kg/ha -
Isolated plains: 15 kg/ha - Hilly or mountainous areas: 10 kg/ha

## References

1.  Agricultural Soil Science Journals
2.  Local Agricultural Extension Services
3.  Soil Fertility Management Guidelines
4.  Sustainable Agricultural Production Techniques \`\`\`

I've integrated the additional details from the Italian document,
expanding the vignette to include: 1. Course objectives 2. Detailed soil
sampling methodology 3. Required soil analyses 4. Expanded explanation
of nitrogen balance calculation 5. More comprehensive interpretation
guidelines 6. Natural nitrogen contribution details

```{r}
library(raster)

# Supponiamo che B04 (Red) e B08 (NIR) siano le bande corrispondenti nel tuo raster
red_band <- s2.rast$B04  # Banda rossa
nir_band <- s2.rast$B08  # Banda NIR

# Calcola l'NDVI
ndvi <- (nir_band - red_band) / (nir_band + red_band)

# Visualizza il risultato
plot(ndvi, main = "NDVI")

```

```{r fig.height=5, fig.width=7.5}
library(raster)

# Esempio di raster NDVI
# Carica un esempio di raster NDVI (sostituisci con il tuo file raster)
# Supponiamo che s2.rast contenga il tuo raster NDVI
# Ad esempio: ndvi_raster <- raster::raster("path/to/ndvi_raster.tif")

# Esegui la calibrazione a due punti
n_rate_raster_2pt <- estimate_N_rate_from_calibration_curve(
  raster = ndvi,  # Supponiamo che la banda NDVI sia s2.rast$NDVI
  minN = 40,               # Il tasso minimo di azoto (kg/ha) per il minimo NDVI
  maxN = 60                # Il tasso massimo di azoto (kg/ha) per il massimo NDVI
)

# Esegui la calibrazione a tre punti (fornendo anche il tasso medio)
n_rate_raster_3pt <- estimate_N_rate_from_calibration_curve(
  raster = ndvi,  # Supponiamo che la banda NDVI sia s2.rast$NDVI
  minN = 40,               # Il tasso minimo di azoto (kg/ha) per il minimo NDVI
  meanN = 50,              # Il tasso medio di azoto (kg/ha) per il valore medio di NDVI
  maxN = 60,               # Il tasso massimo di azoto (kg/ha) per il massimo NDVI
  calibration_type = "three-point",  # Usa la calibrazione a tre punti
  plot = TRUE                       # Mostra il grafico della calibrazione
)

# Visualizza il risultato del raster con i tassi di applicazione di azoto
plot(n_rate_raster_2pt, main = "Nitrogen Application Rate (Two-Point Calibration)")
plot(n_rate_raster_3pt, main = "Nitrogen Application Rate (Three-Point Calibration)")

```

