---
title: "Nitrogen Fertilization Management with NFert: A Comprehensive Guide"
author: "Michele Croci, Stefano Amaducci"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Nitrogen Fertilization Management with NFert: A Comprehensive Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The **NFert** R package provides a comprehensive toolkit for calculating nitrogen fertilization requirements in field crops following the Emilia-Romagna Regional Recommendations for integrated production. This vignette demonstrates how to use NFert for nitrogen balance calculations, soil fertility assessment, and precision agriculture applications.

## What is Nitrogen Balance?

Nitrogen balance is a quantitative method that accounts for all nitrogen inputs and outputs in an agricultural system. The general equation used in NFert is:

$$\text{Nitrogen Fertilization (N)} = A - B + C1 + C2 + D - E - F_{org} - G$$

Where:

- **A**: Crop nitrogen requirements (kg/ha)
- **B**: Soil nitrogen supply (sum of b1 and b2) (kg/ha)
- **C1**: Nitrogen leaching loss in winter (kg/ha)
- **C2**: Nitrogen leaching loss in spring (kg/ha)
- **D**: Immobilization and dispersion losses (kg/ha)
- **E**: Nitrogen from previous crop residues (kg/ha)
- **F_{org}**: Nitrogen from organic fertilizers (kg/ha)
- **G**: Natural nitrogen contributions (kg/ha)

## Normative framework: DPI 2025–2026 and FertDPI

NFert implements the **method of the balance** (Allegato 2) of the Disciplinari di Produzione Integrata (DPI) Emilia-Romagna, in line with the *Guida alla Fertilizzazione Minerale e Organica* (DPI 2025) and the regional tool **FertDPI / Fert_Office**. References: DPI Norme Generali (Allegato 2, 3, 4, 6, 7, 9), Reg. reg. 2/2024 and 3/2017, Direttiva Nitrati (91/676/CEE). In **Zone Vulnerabili ai Nitrati (ZVN)** the MAS (limiti di massima applicazione standard) are binding and the 170 kg N/ha/year ceiling from livestock effluents applies at farm level.

**Formula N (Guida DPI 2025):**  
N da apportare = **A** − **B** + **C** + **D** − **E** − **F** − **G**

- **A** = Fabbisogni colturali (assorbimenti unitari × produzione attesa).  
- **B** = Fertilità del suolo: **B1** (mineralizzazione S.O.: coefficiente × S.O. % × coefficiente tempo) + **B2** (azoto pronto: coefficiente × N totale ‰; solo colture con ciclo &lt; 1 anno). Per pluriennali B = B1 soltanto.  
- **C** = Perdite per lisciviazione: **Ca** (autunno-inverno: &lt;150 mm nessuna; 150–250 mm perdita progressiva; &gt;250 mm tutto l’N pronto) + **Cb** (febbraio: 1 kg N/ha ogni 10 mm oltre 250 mm cumulati).  
- **D** = Immobilizzazione e dispersione = B × fattore di correzione *fc* (in funzione di ossigeno e tessitura: es. buona 0,15–0,25, moderata 0,20–0,30, impedita 0,30–0,40).  
- **E** = N da residui della precessione (tabella: cereali paglia asportata −10, interrata −30; medica +80; soia 0; leguminose da granella +40; ecc.).  
- **F** = N da fertilizzazioni organiche anni precedenti (ammendanti 50/30/20 %; Cattle slurry 30/15/10; suino/pollina 15/10/5).  
- **G** = Apporti naturali (precipitazioni; fissazione per leguminose).

NFert computes the balance and subtracts the current-year organic N (Forg) to obtain the mineral N to apply. Fractioning rules (max 100 kg/ha N per intervention for herbaceous crops, 60 kg/ha for tree crops; min 7 days between interventions) and presemina/autumn caps (e.g. max 30 kg/ha presemina, max 40 kg/ha arboree before 15 Oct) are set by DPI and must be respected in the field plan.

**Soil texture (DPI groupings):** The 12 USDA texture classes are grouped into three macro-categories used for efficiency and coefficients: (1) *Tendenzialmente sabbioso* (S, SF, FS), (2) *Franco / Medio impasto* (L, F, FL, FSA, FA), (3) *Tendenzialmente argilloso* (FLA, AS, AL, A). NFert's `tri3()` and `calc_soil_group_and_id_rag()` provide texture and soil group.

**Maximum allowed doses (MAS):** Each crop has a ceiling for N and P₂O₅ that must not be exceeded. Use `get_MAS(crop)` and `check_MAS(crop, N_planned)` to retrieve and check against MAS (see section below).

**Organic N efficiency (DPI):** The share of organic N available to the crop in the year of application depends on fertilizer type, **distribution technique** (efficient / medium / low), and **soil texture**. Efficient = injection, fertigation, incorporation within 4 h; medium = band spreading with quick incorporation; low = surface spreading without incorporation. Ammendanti compostati use a fixed 40% efficiency. NFert's `organic_fertilization()` uses factors by source and frequency; for DPI-specific efficiency by soil and distribution, refer to the FertDPI 'Efficienza' sheet or the regional guide.

## Package Overview

NFert provides the following categories of functions:

1. **Field-scale nitrogen balance** - `N_balance()` plus the ten
   component helpers (crop demand, soil supply, leaching, immobilisation,
   previous-crop residue, organic contributions, natural deposition).
2. **Phosphorus and potassium balance** - `P_balance()`, `K_balance()`.
3. **Distribution plan** - `plan_distribution()` with MAS and 170 kg N
   ZVN checks.
4. **Precision agriculture** - `variable_rate_N()`,
   `estimate_N_rate_from_calibration_curve()` and
   `estimate_N_rate_from_holland_schepers()` for VI-based VRT;
   `build_strip_prescription()` for machine-width strips along an A-B
   line; `export_prescription()` / `export_prescription_all()` for
   Shapefile, GeoJSON, KML, GeoPackage, John Deere, Trimble and ISOXML
   TASKDATA formats.
5. **Remote-sensing diagnosis** - `compute_vi()` (six VIs),
   `compute_NNI()` / `compute_NNI_from_S2()` / `crop_params_NNI()`
   for the GPR-based pipeline, `nni_from_vi_empirical()` for the
   single-VI regression path, `estimate_biophysical()` for canopy
   trait retrieval.
6. **Farm-level workflow** - `farm_balance()` iterates `N_balance()`
   over every feature of a GeoJSON / Shapefile and returns an
   enriched `sf` layer ready for export.
7. **Shiny application** - `run_app()` opens a six-tab web interface
   with interactive forms, leaflet maps and one-click export.

# Installation and Setup

```{r installation}
# If not already installed
# install.packages("NFert")

# Load the package
library(NFert)

# For precision agriculture functions, also load raster
library(raster)
```

# Getting Started: A Simple Example

Let's start with a basic example calculating the nitrogen balance for a wheat crop:

```{r basic-example}
# Calculate nitrogen balance for winter wheat
wheat_balance <- N_balance(
  expected_yield_tons_ha = 8,
  crop = "Soft wheat FF - strong (grain)",
  ccp = "Autumn-winter crop <150 days",
  sand = 30,
  clay = 25,
  Ntot = 1.2,
  SOM = 2.0,
  CN = 9,
  oxygen_availability = "Normal",
  winter_rain = 160,
  start_spring_rain = 40,
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  location = "Plain adjacent to urbanized areas",
  forg_quantity = 100
)

# Display results
print(wheat_balance)

# Calculate final nitrogen requirement
required_N <- calculate_N_fertilization(wheat_balance)
cat("\nRequired nitrogen fertilization:", round(required_N, 2), "kg/ha\n")
```

# Understanding the Nitrogen Balance Components

## Component 1: Crop Nitrogen Demand (A)

The crop nitrogen demand is calculated based on expected yield and crop-specific nitrogen uptake coefficients:

```{r crop-demand}
# Calculate crop nitrogen demand
crop_demand <- calc_crop_N_demand(
  expected_yield_tons_ha = 10,
  crop = "Soft wheat FF - strong (grain)"
)

print(crop_demand)

# Available crops in the database
head(NFert::uptake_table, 10)
```

## Component 2: Soil Fertility (B = b1 + b2)

Soil fertility accounts for both readily available nitrogen (b1) and mineralizable nitrogen from organic matter (b2):

```{r soil-fertility}
# Calculate soil fertility
soil_N <- soil_fertility(
  Ntot = 1.2,
  SOM = 2.5,
  soil.group = "Sandy textures",
  CN = 9,
  ccp = "Autumn-winter crop <150 days"
)

print(soil_N)
cat("Total soil N supply (B):", soil_N$b1 + soil_N$b2, "kg/ha\n")
```

### Determining Soil Group

First, you need to determine the soil group based on texture:

```{r soil-texture}
# Classify soil texture using simplified method
texture_class <- tri3(clay = 20, sand = 35)
cat("Simplified texture class:", texture_class, "\n")

# Get detailed classification and soil properties
soil_props <- calc_soil_group_and_id_rag(clay = 20, sand = 35)
print(soil_props)
```

## Component 3: Leaching Losses (C1, C2)

Leaching losses depend on rainfall and soil drainage properties:

```{r leaching}
# Calculate leaching losses
leaching <- leaching_loss(
  winter_rain = 180,
  start_spring_rain = 45,
  oxygen_availability = "Normal",
  id_rag = 3,
  b1 = 25
)

print(leaching)
cat("Total leaching loss:", leaching$C1 + leaching$C2, "kg/ha\n")
```

## Component 4: Immobilization Loss (D)

Immobilization occurs when soil microorganisms use nitrogen for their own growth:

```{r immobilization}
# Calculate immobilization loss
immobilization <- calc_N_immobilization_loss(
  B = 50,
  oxygen_availability = "Normal",
  id_rag = 3
)

cat("Immobilization loss (D):", immobilization, "kg/ha\n")
```

## Component 5: Previous Crop Residues (E)

Nitrogen contribution from previous crop residues:

```{r residues}
# Nitrogen from previous crop
residue_N <- nitrogen_from_previous_crop_residues(
  previous_crop = "Winter cereals straw removal"
)

cat("Nitrogen from residues (E):", residue_N, "kg/ha\n")

# Available previous crops
head(NFert::e.table, 10)
```

## Component 6: Organic Fertilizer (Forg)

Contribution from organic fertilizers:

```{r organic}
# Calculate organic fertilizer contribution
organic_N <- organic_fertilization(
  source = "Cattle slurry",
  frequency = "every year",
  quantity = 100
)

cat("Organic fertilizer N (Forg):", organic_N, "kg/ha\n")

# Compare different organic sources (names must match `f.table$source`)
sources <- c("Cattle slurry", "Manures", "Pig slurry or poultry manure")
organic_comparison <- data.frame(
  source = sources,
  N_contribution = sapply(sources, function(s) {
    organic_fertilization(source = s, frequency = "every year", quantity = 100)
  })
)
print(organic_comparison)
```

## Component 7: Natural Contribution (G)

Natural nitrogen inputs from atmospheric deposition:

```{r natural}
# Calculate natural nitrogen contribution
natural_N <- natural_contribution(
  location = "Plain adjacent to urbanized areas",
  ccp = "Autumn-winter crop <150 days"
)

cat("Natural contribution (G):", natural_N, "kg/ha\n")

# Available locations
head(NFert::g.table, 10)
```

# Complete Workflow: Maize Production Example

Let's work through a complete example for maize production:

```{r maize-workflow}
# Step 1: Define field parameters
maize_params <- list(
  expected_yield_tons_ha = 12,  # tons/ha
  crop = "Silage maize (class 700)",  # or "Silage maize (class 700)" (alias)
  ccp = "Spring-summer crop 100–130 days",
  clay = 28,
  sand = 32,
  Ntot = 1.5,
  SOM = 2.8,
  CN = 10,
  oxygen_availability = "Normal",
  winter_rain = 165,
  start_spring_rain = 35,
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  location = "Plain adjacent to urbanized areas",
  forg_quantity = 90
)

# Step 2: Calculate complete nitrogen balance
maize_balance <- do.call(N_balance, maize_params)

# Step 3: Display detailed results
cat("=== MAIZE NITROGEN BALANCE ===\n")
print(maize_balance)

# Step 4: Calculate final requirement
maize_N_requirement <- calculate_N_fertilization(maize_balance)
cat("\nFinal N requirement:", round(maize_N_requirement, 2), "kg/ha\n")

# Step 5: Interpretation
cat("\n=== INTERPRETATION ===\n")
cat("Crop demand (A):", round(maize_balance$A, 1), "kg/ha\n")
cat("Soil supply (B):", round(maize_balance$B, 1), "kg/ha\n")
cat("Leaching losses (C1+C2):", round(maize_balance$C1 + maize_balance$C2, 1), "kg/ha\n")
cat("Immobilization (D):", round(maize_balance$D, 1), "kg/ha\n")
cat("Residue contribution (E):", round(maize_balance$E, 1), "kg/ha\n")
cat("Organic fertilizer (Forg):", round(maize_balance$Forg, 1), "kg/ha\n")
cat("Natural contribution (G):", round(maize_balance$G, 1), "kg/ha\n")
```

# Checking against Maximum Allowed Doses (MAS)

DPI (Allegato 9, Reg. reg. 2/2024) sets **maximum allowed doses (MAS)** of N per crop; in ZVN they are binding. The planned dose must not exceed MAS. Use `get_MAS(crop, edition = "2025")` for the ZVN table from the *Guida alla Fertilizzazione* 2025, or `get_MAS(crop, edition = "2026")` (default) for the FertDPI-style table. Use `check_MAS()` to verify a planned N dose.

```{r mas-check}
# List MAS for main crops (default: edition 2026)
get_MAS()

# DPI 2025 Guida (ZVN table) – different values and crops
get_MAS(edition = "2025")

# Get MAS for a specific crop
get_MAS("Soft wheat FF - strong (grain)")
get_MAS("Soft wheat FF - strong (grain)", edition = "2025")  # 180 kg/ha
get_MAS("Grain maize 500-700 (grain)", edition = "2025")            # 280 kg/ha (irriguo)

# Check if a planned N dose is within MAS
check_MAS("Soft wheat FF - strong (grain)", 180)   # OK
check_MAS("Soft wheat FF - strong (grain)", 220)   # Over MAS
check_MAS("Grain maize 500-700 (grain)", 250)             # OK (MAS 260 or 280 depending on edition)

# After calculating requirement, verify against MAS
required_N <- calculate_N_fertilization(wheat_balance)
check_MAS("Soft wheat FF - strong (grain)", round(required_N, 0))
```

Soia: MAS N = 30 kg/ha (ZVN 2025); in case of failed rhizobium, up to 120 kg/ha is allowed. Reductions: −40 kg/ha after prato ≥3 years; −60 kg/ha after medicaio ≥3 years. In ZVN, respect the 170 kg N/ha/year limit from livestock effluents (farm average). All fertilizations must be recorded in field sheets within 7 days (DPI cap. 11).

# Scenario Comparison

Compare different management scenarios for the same field:

```{r scenarios}
# Base scenario parameters
base_params <- list(
  expected_yield_tons_ha = 10,
  crop = "Soft wheat FF - strong (grain)",
  ccp = "Autumn-winter crop <150 days",
  clay = 22,
  sand = 30,
  Ntot = 1.2,
  SOM = 2.0,
  CN = 9,
  oxygen_availability = "Normal",
  winter_rain = 160,
  start_spring_rain = 40,
  location = "Plain adjacent to urbanized areas"
)

# Scenario 1: No organic fertilization
scenario1 <- do.call(N_balance, c(base_params, list(
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "never",
  forg_quantity = 0
)))
N1 <- calculate_N_fertilization(scenario1)

# Scenario 2: With organic fertilization (80 m³/ha)
scenario2 <- do.call(N_balance, c(base_params, list(
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  forg_quantity = 80
)))
N2 <- calculate_N_fertilization(scenario2)

# Scenario 3: Higher organic rate (150 m³/ha)
scenario3 <- do.call(N_balance, c(base_params, list(
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  forg_quantity = 150
)))
N3 <- calculate_N_fertilization(scenario3)

# Compare scenarios
comparison <- data.frame(
  Scenario = c("No organic", "80 m³/ha organic", "150 m³/ha organic"),
  Organic_N = c(scenario1$Forg, scenario2$Forg, scenario3$Forg),
  N_requirement = c(N1, N2, N3),
  N_saved = c(0, N1 - N2, N1 - N3)
)

print(comparison)
cat("\nNitrogen saved with organic fertilization:\n")
cat("Scenario 2:", round(N1 - N2, 2), "kg/ha (", round((N1-N2)/N1*100, 1), "%)\n")
cat("Scenario 3:", round(N1 - N3, 2), "kg/ha (", round((N1-N3)/N1*100, 1), "%)\n")
```

# Precision Agriculture: NDVI-Based Variable Rate Application

NFert supports precision agriculture through NDVI-based nitrogen rate estimation. Below, **maps** show example rasters (synthetic field); in practice you replace `ndvi_raster` with your Sentinel‑2 or UAV NDVI layer.

## Method 1: Calibration Curve

```{r precision-ag-calibration, fig.width=10, fig.height=4}
# Create example NDVI raster
set.seed(42)
ndvi_raster <- raster(nrows=20, ncols=20, xmn=0, xmx=100, ymn=0, ymx=100)
values(ndvi_raster) <- runif(ncell(ndvi_raster), 0.35, 0.85)

# Two-point calibration
n_rate_2pt <- estimate_N_rate_from_calibration_curve(
  raster = ndvi_raster,
  minN = 40,  # N rate for max NDVI (healthy areas)
  maxN = 80,  # N rate for min NDVI (areas needing more N)
  calibration_type = "two-point",
  plot = FALSE
)

# Three-point calibration (more accurate)
n_rate_3pt <- estimate_N_rate_from_calibration_curve(
  raster = ndvi_raster,
  minN = 40,
  meanN = 60,  # N rate for mean NDVI
  maxN = 80,
  calibration_type = "three-point",
  plot = FALSE
)

# Compare results
cat("Two-point calibration - Mean N rate:", 
    round(cellStats(n_rate_2pt, mean), 2), "kg/ha\n")
cat("Three-point calibration - Mean N rate:", 
    round(cellStats(n_rate_3pt, mean), 2), "kg/ha\n")

# Spatial maps: NDVI and recommended N rate (variable-rate prescription)
op <- par(no.readonly = TRUE)
par(mfrow = c(1, 3), mar = c(2, 2, 2.5, 1))
raster::plot(ndvi_raster, main = "NDVI (example)", axes = FALSE,
             col = grDevices::hcl.colors(100, "Greens", rev = TRUE))
raster::plot(n_rate_2pt, main = "N rate — 2-point (kg/ha)", axes = FALSE,
             col = grDevices::hcl.colors(100, "YlOrRd", rev = TRUE))
raster::plot(n_rate_3pt, main = "N rate — 3-point (kg/ha)", axes = FALSE,
             col = grDevices::hcl.colors(100, "YlOrRd", rev = TRUE))
par(op)
```

## Method 2: Holland & Schepers Algorithm

```{r precision-ag-holland-schepers, fig.width=10, fig.height=4}
# Holland & Schepers method
hs_result <- estimate_N_rate_from_holland_schepers(
  ndvi_raster = ndvi_raster,
  base_N_rate = 60,  # Base application rate
  plot = FALSE
)

cat("Holland & Schepers - Mean N rate:", 
    round(cellStats(hs_result$dose_raster, mean), 2), "kg/ha\n")
cat("Sufficiency Index range:", 
    round(cellStats(hs_result$sufficiency_index_raster, min), 2), "to",
    round(cellStats(hs_result$sufficiency_index_raster, max), 2), "\n")

# Maps: NDVI, recommended N dose, sufficiency index (SI)
op2 <- par(no.readonly = TRUE)
par(mfrow = c(1, 3), mar = c(2, 2, 2.5, 1))
raster::plot(ndvi_raster, main = "NDVI", axes = FALSE,
             col = grDevices::hcl.colors(100, "Greens", rev = TRUE))
raster::plot(hs_result$dose_raster, main = "H&S N dose (kg/ha)", axes = FALSE,
             col = grDevices::hcl.colors(100, "YlOrBr", rev = TRUE))
raster::plot(hs_result$sufficiency_index_raster, main = "Sufficiency index", axes = FALSE,
             col = grDevices::hcl.colors(100, "Blues", rev = TRUE))
par(op2)

# Export maps for farm management systems
# writeRaster(hs_result$dose_raster, "N_application_map.tif", 
#             format="GTiff", overwrite=TRUE)
```

# Batch Processing Multiple Fields

For agricultural consultants managing multiple fields:

```{r batch-processing}
# Create field data
fields_data <- data.frame(
  field_id = c("Field_A", "Field_B", "Field_C"),
  expected_yield = c(9, 11, 8),
  crop = c("Soft wheat FF - strong (grain)", "Silage maize (class 700)",
           "Soft wheat FF - strong (grain)"),
  ccp = c("Autumn-winter crop <150 days", "Spring-summer crop 100–130 days",
          "Autumn-winter crop <150 days"),
  clay = c(20, 30, 18),
  sand = c(32, 28, 35),
  Ntot = c(1.1, 1.6, 1.0),
  SOM = c(1.9, 3.0, 1.7),
  CN = c(8, 11, 9),
  winter_rain = c(150, 180, 140),
  prev_crop = c("Winter cereals straw removal", "Winter cereals straw removal",
                "Legumes"),
  organic_quantity = c(0, 100, 80)
)

# Function to process one field
process_field <- function(row) {
  N_balance(
    expected_yield_tons_ha = row$expected_yield,
    crop = as.character(row$crop),
    ccp = as.character(row$ccp),
    clay = row$clay,
    sand = row$sand,
    Ntot = row$Ntot,
    SOM = row$SOM,
    CN = row$CN,
    winter_rain = row$winter_rain,
    start_spring_rain = 40,
    prev_crop = as.character(row$prev_crop),
    source = "Cattle slurry",
    fertorg_frequency = ifelse(row$organic_quantity > 0, "every year", "never"),
    forg_quantity = row$organic_quantity,
    oxygen_availability = "Normal",
    location = "Plain adjacent to urbanized areas"
  )
}

# Process all fields
field_results <- lapply(seq_len(nrow(fields_data)), function(i) {
  process_field(fields_data[i, ])
})

# Calculate N requirements
fields_data$N_requirement <- sapply(field_results, function(x) {
  calculate_N_fertilization(x)
})

# Add field areas and calculate totals
fields_data$area_ha <- c(15, 20, 12)
fields_data$total_N_kg <- fields_data$N_requirement * fields_data$area_ha

# Summary table
summary_table <- fields_data[, c("field_id", "crop", "area_ha", 
                                  "N_requirement", "total_N_kg")]
names(summary_table)[4] <- "N_rate_kg_ha"
print(summary_table)

cat("\nTotal nitrogen required for all fields:", 
    sum(summary_table$total_N_kg), "kg\n")
```

# Advanced: Sensitivity Analysis

Understand how different factors affect nitrogen requirements:

```{r sensitivity}
# Base scenario
base_scenario <- N_balance(
  expected_yield_tons_ha = 10,
  crop = "Soft wheat FF - strong (grain)",
  ccp = "Autumn-winter crop <150 days",
  clay = 22, sand = 30,
  Ntot = 1.2, SOM = 2.0, CN = 9,
  winter_rain = 160, start_spring_rain = 40,
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  forg_quantity = 100,
  oxygen_availability = "Normal",
  location = "Plain adjacent to urbanized areas"
)

base_N <- calculate_N_fertilization(base_scenario)

# Sensitivity to SOM
SOM_values <- c(1.5, 2.0, 2.5, 3.0, 3.5)
SOM_impact <- sapply(SOM_values, function(som) {
  sc <- N_balance(
    expected_yield_tons_ha = 10,
    crop = "Soft wheat FF - strong (grain)",
    ccp = "Autumn-winter crop <150 days",
    clay = 22, sand = 30,
    Ntot = 1.2, SOM = som, CN = 9,
    winter_rain = 160, start_spring_rain = 40,
    prev_crop = "Winter cereals straw removal",
    source = "Cattle slurry",
    fertorg_frequency = "every year",
    forg_quantity = 100,
    oxygen_availability = "Normal",
    location = "Plain adjacent to urbanized areas"
  )
  calculate_N_fertilization(sc)
})

# Sensitivity to winter rainfall
rainfall_values <- c(100, 130, 160, 190, 220)
rainfall_impact <- sapply(rainfall_values, function(wr) {
  sc <- N_balance(
    expected_yield_tons_ha = 10,
    crop = "Soft wheat FF - strong (grain)",
    ccp = "Autumn-winter crop <150 days",
    clay = 22, sand = 30,
    Ntot = 1.2, SOM = 2.0, CN = 9,
    winter_rain = wr, start_spring_rain = 40,
    prev_crop = "Winter cereals straw removal",
    source = "Cattle slurry",
    fertorg_frequency = "every year",
    forg_quantity = 100,
    oxygen_availability = "Normal",
    location = "Plain adjacent to urbanized areas"
  )
  calculate_N_fertilization(sc)
})

# Display results
sensitivity_results <- data.frame(
  Parameter = c(rep("SOM (%)", length(SOM_values)),
                rep("Winter rainfall (mm)", length(rainfall_values))),
  Value = c(SOM_values, rainfall_values),
  N_requirement = c(SOM_impact, rainfall_impact),
  Change_from_base = c(SOM_impact - base_N, rainfall_impact - base_N)
)

print(sensitivity_results)

cat("\nKey insights:\n")
cat("- SOM increase from 1.5% to 3.5% changes N requirement by",
    round(max(SOM_impact) - min(SOM_impact), 1), "kg/ha\n")
cat("- Rainfall increase from 100mm to 220mm changes N requirement by",
    round(max(rainfall_impact) - min(rainfall_impact), 1), "kg/ha\n")
```

# Best Practices and Tips

## 1. Respect DPI and MAS

- **MAS**: Always check that the planned N (and P₂O₅ if you use mineral P) does not exceed the maximum allowed dose for the crop: `check_MAS(crop, N_planned)`.
- **ZVN**: In Nitrate Vulnerable Zones, respect the 170 kg N/ha/year ceiling from livestock effluents (farm level).
- **Fertilization plan**: DPI requires a written plan (scadenze: erbacee entro 28 feb, orticole entro 15 apr, arboree/sementiere entro 15 apr; versione definitiva before harvest). Register each intervention in the field sheets within 7 days.

## 2. Input Data Quality

- **Soil sampling**: Sample at optimal times (autumn, 2 months after fertilization)
- **Sample depth**: 0-30 cm is standard for nitrogen balance calculations
- **Number of samples**: 15-20 points per field for representative results
- **Laboratory analysis**: Use certified laboratories for accurate results

## 3. Crop Selection

Make sure to use the exact crop name as it appears in the database:

```{r crop-selection}
# Check available crops
available_crops <- NFert::uptake_table$crop
cat("Available crops (first 10):\n")
print(head(available_crops, 10))
```

## 4. Parameter Validation

The `N_balance()` function includes input validation, but you should verify:

- Sand + Clay ≤ 100%
- All percentages are between 0-100
- Expected yield is positive
- Crop name matches database exactly

## 5. Interpreting Results

- **Negative N requirement**: Soil supply exceeds crop demand - no fertilization needed
- **Low N requirement (<50 kg/ha)**: Soil is fertile, minimal supplementation needed
- **High N requirement (>150 kg/ha)**: Consider organic amendments or soil improvement

## 6. Seasonal Adjustments

- Recalculate after significant rainfall events (affects leaching)
- Monitor soil moisture (affects mineralization rates)
- Adjust for seasonal variations in organic matter decomposition

# Troubleshooting Common Issues

## Issue 1: Crop not found

```{r troubleshooting-1}
# Error: Crop not found
# Solution: Check exact spelling and match with database
tryCatch({
  calc_crop_N_demand(crop = "Wheat")  # Wrong name
}, error = function(e) {
  cat("Error:", e$message, "\n")
  cat("Solution: Use 'Soft wheat FF - strong (grain)' instead\n")
})
```

## Issue 2: Invalid soil texture

```{r troubleshooting-2}
# Error: Invalid clay and sand combination
# Solution: Ensure clay + sand ≤ 100
tryCatch({
  tri3(clay = 60, sand = 50)  # Sum exceeds 100
}, error = function(e) {
  cat("Error:", e$message, "\n")
  cat("Solution: Check that clay + sand ≤ 100%\n")
})
```

## Issue 3: NA values in results

```{r troubleshooting-3}
# If you get NA values, check:
# 1. Location name matches g.table
# 2. CCP matches coef_time
# 3. Previous crop matches e.table
# 4. Oxygen availability matches ca.table

# The function will warn you about NA components
result <- N_balance(
  expected_yield_tons_ha = 10,
  crop = "Soft wheat FF - strong (grain)",
  ccp = "Autumn-winter crop <150 days",
  sand = 30, clay = 25,
  # ... other parameters
)
# Check for warnings about missing data
```

# Additional Resources

## Internal Data Tables

All internal datasets are documented and accessible:

```{r data-tables}
# View available datasets
cat("Available internal datasets:\n")
cat("- uptake_table: Crop nitrogen uptake coefficients\n")
cat("- e.table: Previous crop residue values\n")
cat("- f.table: Organic fertilizer factors\n")
cat("- g.table: Natural deposition rates\n")
cat("- soil.table: Soil classification table\n")
cat("- tri2.table, tri3.table: Soil texture matrices\n")
cat("- ca.table, cb.table: Leaching and immobilization factors\n")
cat("- coefN_readily, coefN_mineralised: Soil fertility coefficients\n")
cat("- coef_time: Time adjustment factors\n")

# Access any dataset
head(NFert::uptake_table, 5)
```

## Getting Help

- **Package documentation**: `help(package = "NFert")`
- **Function help**: `?N_balance` or `?calc_crop_N_demand`
- **GitHub issues**: Report bugs at https://github.com/mcroci/NFert/issues
- **Email**: Contact michele.croci@unicatt.it

# References

1. Disciplinari di Produzione Integrata (DPI) 2025 – Regione Emilia-Romagna, Direzione Generale Agricoltura, Caccia e Pesca. Norme Generali, Allegati 2, 3, 4, 6, 7, 9.
2. Guida alla Fertilizzazione Minerale e Organica – N, P, K per colture erbacee e arboree. DPI 2025, Emilia-Romagna.
3. Regolamento regionale 2/2024 – Limiti MAS (Allegato 9). Regolamento regionale 3/2017 – Utilizzazione agronomica degli effluenti e del digestato.
4. Direttiva Nitrati 91/676/CEE – Zone Vulnerabili ai Nitrati (170 kg N/ha/anno da effluenti zootecnici; soglie minime efficienza Reg. UE 2021/2115).
5. FertDPI / Fert_Office – strumento regionale per il piano di fertilizzazione (DPI 2026).
6. Precision Agriculture Best Practices (International Society of Precision Agriculture).

---

**Note**: The calculations in NFert follow regional guidelines for Emilia-Romagna, Italy. For other regions, coefficients and parameters may need adjustment based on local conditions and regulations.
