---
title: "N, P and K fertilization with NFert — DPI Emilia-Romagna 2026"
author: "Michele Croci, Manuele Ragazzi, Giorgio Impollonia, Stefano Amaducci"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{NFert PK and Distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 6,
  fig.height = 4,
  fig.align  = "center",
  dpi        = 120
)
```

# Introduction

This vignette shows how to use **NFert 0.4.0** to compute a full nitrogen,
phosphorus and potassium fertilisation plan following the
*Disciplinari di Produzione Integrata* (DPI) Emilia-Romagna 2026. The
package reproduces the logic of the official regional tool
**Fert_Office v1.26** (Febbraio 2026, Regione Emilia-Romagna) and makes it
scriptable in R.

The package implements **two complementary methods**:

1. **Metodo del bilancio (Allegato 2)**: full nutrient mass balance with
   crop demand, soil supply, losses, previous-crop effects and organic
   contributions. Recommended when detailed soil and climate data are
   available.
2. **Scheda a dose Standard**: simplified per-crop dose with user-selectable
   decrement and increment factors, capped by MAS (Massimali Ammessi di
   Simulazione). Recommended when soil analyses are limited.

The package also provides a **fertilisation distribution plan** and
**end-of-cycle soil P and K estimation** to close the agronomic loop between
planning and recording.

## Normative reference

- **DPI Emilia-Romagna 2026** — Norme Generali, Allegato 2 (metodo del
  bilancio), Allegato 9 (massimali), *Guida alla Fertilizzazione Minerale
  e Organica* (N, P, K).
- **Reg. reg. 2/2024** — limiti MAS.
- **Reg. reg. 3/2017** — utilizzazione agronomica degli effluenti e del
  digestato.
- **Direttiva Nitrati 91/676/CEE** — in ZVN limite di 170 kg N/ha/anno da
  effluenti zootecnici.
- **Reg. UE 2021/2115** — soglie minime di efficienza.

# Naming conventions

NFert mixes English and Italian terminology by design:

- **Function names**: English snake_case (e.g. `N_balance`, `P_balance`,
  `plan_distribution`, `classify_pH`). Three exceptions retain the Italian
  DPI canonical name: `scheda_N`, `scheda_PK`, `get_MAS` / `check_MAS`.
- **Parameters**: English snake_case throughout.
- **Class strings** (e.g. "molto bassa", "media", "Sabbiosi") are kept in
  Italian because they are DPI canonical labels.
- **Soil-group labels** are accepted in any of three conventions:
  Italian plural ("Sabbiosi", "Medio impasto", "Argillosi e limosi"),
  Italian singular ("Sabbioso", "Franco", "Argilloso") or English
  ("Sandy textures", "Loamy textures", "Clay textures"). The
  `normalise_soil_group()` helper canonicalises any of them.

```{r naming}
library(NFert)
normalise_soil_group("Loamy textures")
normalise_soil_group("Franco")
resolve_id_rag(clay = 18.5, sand = 15.5)
```

# Setup

```{r setup, message = FALSE}
# Crop and yield
crop  <- "Durum wheat (whole plant)"
yield <- 6   # t/ha

# Soil (sample: Sabbia 15.5%, Argilla 18.5% -> FL, Medio impasto)
clay   <- 18.5
sand   <- 15.5
Ntot   <- 1.5     # g/kg
SOM    <- 2       # %
CN     <- 7.73
pH     <- 8.3
caco3_tot <- 15   # %
caco3_att <- 6.8  # %
olsen_P   <- 15   # ppm P2O5
K         <- 150  # ppm K2O

# Climate
winter_rain       <- 150
start_spring_rain <- 0

# Management
prev_crop         <- "Maize stalks removed"
location          <- "Plain adjacent to urbanized areas"
soil_seeding      <- "traditional"
greenhouse        <- FALSE
```

The scenario above reproduces the sample cell of `Fert_Office v1.26 !
Inserimento` and is used as the end-to-end benchmark of the package.

# Soil characterisation

NFert classifies the soil along the DPI 2026 dimensions. The 3-class texture
grouping (Sabbiosi / Medio impasto / Argillosi e limosi) drives every
subsequent coefficient choice (B, C, D, efficiency, soil weight, etc.).

```{r soil-class}
# USDA class + DPI 3-group
soil_props <- calc_soil_group_and_id_rag(clay = clay, sand = sand)
soil_props

# pH, calcare, SO, CSC
classify_pH(pH)
classify_carbonate_tot(caco3_tot)
classify_carbonate_att(caco3_att)
som_cls <- classify_SOM(SOM = SOM, soil_group = "Medio impasto")
som_cls

# Maximum annual SO input (for ammendanti)
max_SO_input(som_cls$class)
```

# Nitrogen balance (Allegato 2)

The DPI 2026 formula is:

$$
N_{fert} = A - B + C_1 + C_2 + D - E - F - F_{org} - G
$$

where each term is computed by a dedicated function:

| Term | Function | Meaning |
|---|---|---|
| A | `calc_crop_N_demand()` | Crop N demand (kg/ha) |
| B = b1 + b2 | `soil_fertility()` | b1 = N pronto from N total, b2 = N mineralised from SOM |
| C1, C2 | `leaching_loss()` | Autumn-winter and early-spring leaching |
| D | `calc_N_immobilization_loss()` | Microbial immobilisation |
| E | `nitrogen_from_previous_crop_residues()` | Previous-crop residual N |
| F | `organic_previous_years_N()` | Residual N from previous years' organic |
| Forg | `organic_fertilization()` | Current-year organic N, DPI efficiency |
| G | `natural_contribution()` | Atmospheric deposition |

## Full computation

```{r n-balance}
n_bal <- N_balance(
  expected_yield_tons_ha = yield,
  crop = crop,
  ccp  = "Spring-summer crop 100-130 days",
  sand = sand, clay = clay,
  Ntot = Ntot, SOM = SOM, CN = CN,
  oxygen_availability = "Normal",
  winter_rain = winter_rain,
  start_spring_rain = start_spring_rain,
  prev_crop = prev_crop,
  source = "None",
  fertorg_frequency = "every year",
  location = location,
  forg_quantity = 0,
  soil_seeding = soil_seeding,
  greenhouse = greenhouse,
  E_to_D = TRUE        # fold negative E into D, as in foglio Bilancio
)

n_bal

# Final dose
n_dose <- calculate_N_fertilization(n_bal)
n_dose
```

```{r fig-n-balance, fig.width=7, fig.height=5, out.width="100%"}
# Signed terms in N_fert = A - B + C1 + C2 + D - E - F - Forg - G (kg N/ha)
signed <- c(
  A = n_bal$A, B = -n_bal$B, C1 = n_bal$C1, C2 = n_bal$C2, D = n_bal$D,
  E = -n_bal$E, F = -n_bal$F, Forg = -n_bal$Forg, G = -n_bal$G
)
op <- par(mar = c(7, 4, 3, 1))
cols <- grDevices::hcl.colors(length(signed), "BluYl", alpha = 0.9)
barplot(
  as.numeric(signed), names.arg = names(signed), las = 2, col = cols,
  border = NA, ylab = "kg N/ha",
  main = "N balance terms (signed toward mineral N need)"
)
abline(h = 0, lty = 2, col = "gray40")
par(op)
```

The result matches the Fert_Office `Bilancio` sheet: A = 186.6, B = 73.84,
D = 39.536 (includes `|E|` = 10), G = 10.05, and
**N required ≈ 142.25 kg/ha**.

## MAS verification (Massimali Ammessi di Simulazione)

```{r mas}
get_MAS(crop = crop)
check_MAS(crop, n_dose)
```

```{r fig-mas, fig.width=5.5, fig.height=4.5, out.width="85%"}
mas_row <- get_MAS(crop = crop)
mas_n <- mas_row$mas_N[1]
vals <- c(`Planned N` = n_dose, `MAS N` = mas_n)
bp <- barplot(
  vals, col = grDevices::hcl.colors(2, "Peach", rev = TRUE), border = NA,
  ylab = "kg N/ha", main = "Planned N dose vs maximum allowed (MAS)"
)
text(bp, vals + 0.04 * max(vals, na.rm = TRUE), labels = round(vals, 1), pos = 3, cex = 0.9)
```

## Component functions (pedagogical use)

Every balance term is also exposed as a standalone function for teaching or
advanced scenario analysis:

```{r n-components}
demand <- calc_crop_N_demand(expected_yield_tons_ha = yield, crop = crop)
demand

fertility <- soil_fertility(Ntot = Ntot, SOM = SOM,
                            soil.group = "Loamy textures", CN = CN,
                            ccp = "Spring-summer crop 100-130 days",
                            soil_seeding = soil_seeding)
fertility

leach <- leaching_loss(winter_rain = winter_rain,
                       start_spring_rain = start_spring_rain,
                       oxygen_availability = "Normal",
                       id_rag = soil_props$id_rag,
                       b1 = fertility$b1)
leach

imm <- calc_N_immobilization_loss(B = fertility$b1 + fertility$b2,
                                  oxygen_availability = "Normal",
                                  id_rag = soil_props$id_rag,
                                  greenhouse = greenhouse,
                                  E_residual = -10)   # stocchi asportati
imm

natural_contribution(location = location,
                     ccp = "Spring-summer crop 100-130 days")
```

# Phosphorus balance

The phosphorus balance follows the DPI Allegato 2 logic coded in foglio
`Gri_P`:

$$
P_2O_5 = A_{asp} + B_{1,arric} + A_2 - B_{2,rid}
$$

with the *strategia* depending on the soil Olsen class:

- **Molto bassa / bassa**: `Arricchimento` — add $B_1 = (22.9 -
  P_2O_5^{ppm}) \times w_{30cm}/1000 \times f_{imm}$ where $f_{imm}=1.6$.
- **Media / elevata**: `Mantenimento` — apply A = asportazione.
- **Molto elevata**: `Riduzione` — A = 0.

## Olsen classification

```{r p-class}
p_cls <- classify_P_olsen(value = olsen_P, unit = "P2O5")
p_cls
```

## P balance

```{r p-balance}
p_bal <- P_balance(
  expected_yield_tons_ha = yield, crop = crop,
  olsen_value = olsen_P, olsen_unit = "P2O5",
  clay = clay, sand = sand
)
p_bal
```

In the benchmark scenario (Olsen P = 15 ppm P2O5) the class is "bassa"
→ Arricchimento, so:

- $A = 63.6$ kg P2O5/ha (asportazione)
- $B_1 = (22.9 - 15) \times 3900/1000 \times 1.6 = 49.3$ kg P2O5/ha
- Total **≈ 112.9 kg P2O5/ha**

This matches cell `Bilancio!I33` of Fert_Office.

# Potassium balance

The potassium balance extends the formula with a leaching term **H** that
depends on clay content:

$$
K_2O = A_{asp} + H + B_{1,arric} + A_2 - B_{2,rid}
$$

Classes for K are defined per DPI texture group (`Sabbiosi`,
`Medio impasto`, `Argillosi e limosi`).

## K classification and leaching

```{r k-class}
k_cls <- classify_K(value = K, unit = "K2O", soil_group = "Medio impasto")
k_cls

# Leaching by clay (step function):
sapply(c(2, 10, 18.5, 30), K_leaching_by_clay)
```

```{r fig-k-leach, fig.width=7, fig.height=4.5, out.width="100%"}
clay_seq <- seq(0, 45, length.out = 300)
h_seq <- vapply(clay_seq, K_leaching_by_clay, numeric(1))
plot(
  clay_seq, h_seq, type = "s", lwd = 2,
  col = grDevices::hcl.colors(1, "Teal"),
  xlab = "Clay (%)",
  ylab = expression("H — K leaching coefficient (kg K"[2]*"O/ha)"),
  main = "K balance: clay-dependent leaching term (DPI step function)"
)
rug(c(2, 10, 18.5, 30), col = "gray55")
points(
  c(2, 10, 18.5, 30), sapply(c(2, 10, 18.5, 30), K_leaching_by_clay),
  pch = 21, bg = "coral", col = "white", cex = 1.2
)
```

## K balance

```{r k-balance}
k_bal <- K_balance(
  expected_yield_tons_ha = yield, crop = crop,
  k_value = K, k_unit = "K2O",
  clay = clay, sand = sand
)
k_bal
```

```{r fig-npk, fig.width=6.5, fig.height=4.5, out.width="100%"}
npk <- c(
  `N (mineral)` = n_dose,
  `P2O5` = p_bal$P2O5_required,
  `K2O` = k_bal$K2O_required
)
barplot(
  npk, col = grDevices::hcl.colors(3, "Dark 3", alpha = 0.92), border = NA,
  ylab = "kg/ha",
  main = "Recommended N, P2O5 and K2O inputs (benchmark scenario)"
)
```

Scenario: K = 150 ppm K2O is "media" in Medio impasto → Mantenimento.
- $A = 119.4$ kg K2O/ha
- $H = 20$ (clay 18.5% → step "15.1-25.1 → 20")
- Total **= 139.4 kg K2O/ha**

This matches cell `Gri_K!F30` of Fert_Office (A + H).

# Scheda a dose Standard

For contexts with limited soil analyses, DPI 2026 provides a standard-dose
method. NFert mirrors the flag catalogue of `Fert_Office` `Scheda_N` and
`Scheda_PK`.

## Scheda N

The N standard dose is modified by five decrements and five increments,
then capped by MAS.

```{r scheda-n}
# Grano duro with buried straw + compacted / no-till
scheda_N(
  crop = crop,
  increments = c(straw_burial = 30, compacted_no_till = 10)
)

# Alfalfa after, yield low
scheda_N(
  crop = crop,
  decrements = c(after_alfalfa_meadow = 60, yield_low = 20),
  increments = c(low_SOM = 20)
)
```

When the cumulated dose overshoots `max_N_dose`, the function caps it and
flags `mas_exceeded = TRUE`:

```{r scheda-n-cap}
scheda_N(
  crop = crop,
  increments = c(straw_burial = 100, compacted_no_till = 100, low_SOM = 50)
)
```

## Scheda PK

For P and K the base dose depends on the initial soil dotation class:

```{r scheda-pk}
scheda_PK(
  crop = crop,
  soil_P_class = "normal",
  soil_K_class = "normal"
)

# Example: enrichment under low P dotation
scheda_PK(
  crop = crop,
  soil_P_class = "low",
  soil_K_class = "normal",
  P_increments = c(ristoppio = 20)
)
```

# Fertilisation distribution plan

Once the required N, P, K doses are known, the plan distributes them across
epochs and methods. `plan_distribution()` accepts:

- A list of organic applications (matrix name from `organic_fertilizers.table`,
  quantity in t/ha, year, modality-epoch id, optional efficiency level).
- A list of mineral applications (name from `mineral_fertilizers.table`, quantity in
  q/ha, modality-epoch).
- Optional target doses (for Eccesso/Deficit alerts) and a ZVN flag.

```{r plan}
plan <- plan_distribution(
  soil_group = "Medio impasto",
  n_balance  = n_dose,
  p_balance  = p_bal$P2O5_required,
  k_balance  = k_bal$K2O_required,
  organic_rows = list(
    list(fertilizer    = "letame bovino",
         quantity_t_ha = 20,
         year          = 2025,
         modality_epoch = 1,
         level         = "media")
  ),
  mineral_rows = list(
    list(fertilizer       = "UREA AGRICOLA PRIL.46%",
         quantity_q_ha = 3,
         modality_epoch = 11)
  ),
  zvn = TRUE
)

# Row-by-row plan
plan$rows
# Cumulated useful nutrients
plan$totals
# Target vs delivered alerts
plan$alerts
# ZVN 170 kg N/ha check
plan$zvn_alert
```

## Choosing a distribution modality

Modalities are catalogued in `distribution_modalities.table` and their compatibility
with crop cycles in `cycle_modality.table`:

```{r mod}
head(NFert::distribution_modalities.table, 12)
```

## Fertiliser sources

Available organic matrices and mineral products:

```{r fert-sources}
# Organic: 21 matrices with their N, P, K and dry-matter titres
head(NFert::organic_fertilizers.table[, c("fertilizer", "avg_dm", "avg_N",
                               "avg_P2O5", "avg_K2O", "fully_zootec")], 10)

# Mineral: 146 commercial fertilisers
head(NFert::mineral_fertilizers.table, 10)
```

## DPI 2026 organic N efficiency

The N efficiency of organic fertilisers depends on the soil group (3
classes), the dose level (bassa / media / alta), the sector (avi, bov, sui,
dig_tq, dig_sui, dig_avi, dig_chi, fanghi) and the modality level. The full
matrix has 220 entries:

```{r eff}
# Medium impasto, dose media (125-249 kg N/ha), bovino:
NFert::efficiency.table[
  NFert::efficiency.table$ID_Rag == 2 &
    NFert::efficiency.table$ID_Liv == 2 &
    NFert::efficiency.table$sector_id == "bov",
]
```

# End-of-cycle soil dotation

After the plan is executed, NFert estimates the residual soil P and K in
ppm, usable as input for the following year's calculation:

```{r end}
P_end <- estimate_soil_P_end_of_cycle(
  P2O5_start_ppm = olsen_P,
  P2O5_applied   = plan$totals["P2O5_useful"],
  P2O5_removed   = p_bal$A_fabbisogno,
  soil_group     = "Medio impasto"
)
P_end

K_end <- estimate_soil_K_end_of_cycle(
  K2O_start_ppm = K,
  K2O_applied   = plan$totals["K2O_useful"],
  K2O_removed   = k_bal$A_fabbisogno,
  clay_pct      = clay,
  soil_group    = "Medio impasto"
)
K_end
```

# Alternative scenario: maize (forage, irrigated)

A second scenario with different crop, cycle, irrigation and precessione.

```{r maize}
mais_bal <- N_balance(
  expected_yield_tons_ha = 15,
  crop = "Maize silage (class 700)",
  ccp  = "Spring-summer crop 100-130 days",
  sand = 50, clay = 30,
  Ntot = 1.2, SOM = 1.6, CN = 10,
  oxygen_availability = "Normal",
  winter_rain = 260, start_spring_rain = 40,
  prev_crop = "Winter cereals straw removal",
  source = "Cattle slurry",
  fertorg_frequency = "every year",
  location = "Isolated plain",
  forg_quantity = 80,
  soil_seeding = "traditional"
)
mais_bal
calculate_N_fertilization(mais_bal)
```

# Alternative scenario: apple orchard in piena produzione

```{r apple}
apple_n <- N_balance(
  expected_yield_tons_ha = 40,
  crop = "Melo frutti, legno e foglie",
  ccp  = "In production",
  sand = 40, clay = 30,
  Ntot = 1.8, SOM = 2.5, CN = 9,
  oxygen_availability = "Normal",
  winter_rain = 220, start_spring_rain = 30,
  prev_crop = "Undefined",
  source = "None", fertorg_frequency = "every year",
  location = "Hill or mountain",
  forg_quantity = 0
)
apple_n
calculate_N_fertilization(apple_n)

apple_p <- P_balance(expected_yield_tons_ha = 40,
                     crop = "Melo frutti, legno e foglie",
                     olsen_value = 25, olsen_unit = "P2O5",
                     clay = 30, sand = 40)
apple_p

apple_k <- K_balance(expected_yield_tons_ha = 40,
                     crop = "Melo frutti, legno e foglie",
                     k_value = 180, k_unit = "K2O",
                     clay = 30, sand = 40)
apple_k
```

# Precision agriculture: variable-rate from balance + NDVI

NFert bridges the **agronomic balance** (constant-rate dose, kg/ha) with
**NDVI-based variable-rate application**. The integration logic is:

1. Compute the **field-average dose** with `N_balance()` +
   `calculate_N_fertilization()` or `scheda_N()`.
2. Use NDVI as a proxy of crop vigour: low NDVI → more N (rescue),
   high NDVI → less N (already vigorous).
3. Apply one of the two NFert variable-rate algorithms:

   - **Calibration curve** (`estimate_N_rate_from_calibration_curve()`):
     two-point or three-point linear interpolation between observed NDVI
     extremes and user-defined N rates.
   - **Holland & Schepers** (`estimate_N_rate_from_holland_schepers()`):
     sufficiency-index method using the q95 NDVI as reference and a
     base N rate.

The wrapper `variable_rate_N()` glues balance and NDVI in one call,
preserves the field-average dose by rescaling and (optionally) caps each
pixel at the MAS limit.

## Example: variable-rate N for grano duro

```{r prec-ag-setup, eval = requireNamespace("raster", quietly = TRUE)}
library(raster)

# NDVI raster: example from NFert (RasterLayer or multi-band stack; functions use the NDVI layer if named)
data(s2.rast)
ndvi <- s2.rast
ndvi
```

```{r prec-ag-calibration, eval = requireNamespace("raster", quietly = TRUE), fig.width = 6, fig.height = 5}
# Field-average N from the balance
n_dose                 # ~142 kg/ha (computed earlier in this vignette)

# Variable-rate via two-point calibration; envelope of +/-25%
vr_cal <- variable_rate_N(
  ndvi_raster = ndvi,
  n_dose      = n_dose,
  method      = "calibration",
  envelope    = 0.25,
  mas_cap     = 190     # MAS for grano duro (RR 2/2024)
)

vr_cal$mean_kg_ha       # ~ n_dose
vr_cal$min_kg_ha
vr_cal$max_kg_ha

raster::plot(vr_cal$rate_raster,
             main = "Variable rate N (calibration) - kg/ha",
             col  = grDevices::hcl.colors(50, "YlOrBr", rev = TRUE))
```

## Example: Holland & Schepers method

```{r prec-ag-holland, eval = requireNamespace("raster", quietly = TRUE), fig.width = 6, fig.height = 5}
vr_hs <- variable_rate_N(
  ndvi_raster = ndvi,
  n_dose      = n_dose,
  method      = "holland",
  mas_cap     = 190
)

vr_hs$mean_kg_ha
raster::plot(vr_hs$rate_raster,
             main = "Variable rate N (Holland & Schepers) - kg/ha",
             col  = grDevices::hcl.colors(50, "YlOrBr", rev = TRUE))
```

## Comparison and aggregation

```{r prec-ag-compare, eval = requireNamespace("raster", quietly = TRUE)}
# Total N applied to the field (assuming each pixel has the same area)
mean(getValues(vr_cal$rate_raster), na.rm = TRUE)
mean(getValues(vr_hs$rate_raster),  na.rm = TRUE)

# Histogram comparison
oldpar <- par(mfrow = c(1, 2))
hist(getValues(vr_cal$rate_raster), main = "Calibration",
     xlab = "N rate (kg/ha)", col = "steelblue", border = "white")
hist(getValues(vr_hs$rate_raster),  main = "Holland & Schepers",
     xlab = "N rate (kg/ha)", col = "tomato",   border = "white")
par(oldpar)
```

## Linking to the distribution plan

The variable-rate raster can be coupled to a single fertiliser application
(typically the **copertura**) inside `plan_distribution()`: read the mean
rate from the raster and pass it as the mineral row.

```{r prec-ag-plan, eval = requireNamespace("raster", quietly = TRUE)}
mean_rate <- vr_cal$mean_kg_ha
# Quintals/ha of UREA needed to deliver mean_rate (titre 46% N)
q_urea_per_ha <- mean_rate / 46

plan_vr <- plan_distribution(
  soil_group = "Medio impasto",
  n_balance  = n_dose, p_balance = 0, k_balance = 0,
  organic_rows = list(),
  mineral_rows = list(
    list(fertilizer       = "UREA AGRICOLA PRIL.46%",
         quantity_q_ha = q_urea_per_ha,
         modality_epoch = 11)            # In copertura - pieno accestimento
  )
)
plan_vr$rows
plan_vr$totals
```

## Three-point calibration (advanced)

For more refined NDVI-rate relationships you can call the underlying
estimator directly, providing a `meanN` for the average NDVI:

```{r prec-ag-3pt, eval = requireNamespace("raster", quietly = TRUE)}
n_rate_3pt <- estimate_N_rate_from_calibration_curve(
  raster = ndvi,
  minN = 100, meanN = 140, maxN = 180,
  calibration_type = "three-point"
)
raster::cellStats(n_rate_3pt, summary)
```

# Reference tables

NFert exposes 33 internal tables; the most useful in day-to-day work:

```{r tables}
# Crop master list
head(NFert::uptake_table[, c("crop_id", "crop", "N", "P2O5", "K2O",
                              "reference_yield")], 5)

# MAS (massimali) 2026
head(NFert::mas.table[, c("crop_id", "crop", "MAS_N", "standard_N",
                           "max_N_dose")], 5)

# Gri_P and Gri_K classes
NFert::p_availability.table
NFert::k_availability.table

# Texture grouping with soil weights
NFert::texture_groups.table
```

# Notes and limitations

- The package implements the Italian DPI Emilia-Romagna specification.
  Coefficients and thresholds differ from other Italian regions or national
  standards.
- The `plan_distribution()` function checks totals vs balance targets and
  the ZVN 170 kg N/ha limit, but does not yet check MAS per single
  application nor the specific modality-cycle DPI constraints
  (`cycle_modality.table`). A future release will add these checks.
- The `s2.rast` dataset is a minimal Sentinel-2 NDVI raster used only for
  vignettes and examples; for production use, replace it with your own NDVI
  raster (Sentinel-2, drone, or any other source) at the desired
  resolution.
- `variable_rate_N()` re-scales the raster so the field-average matches the
  agronomic dose; if you instead want to allow the mean to drift (e.g. when
  upper/lower envelopes are explicitly chosen), use the underlying
  `estimate_N_rate_from_calibration_curve()` /
  `estimate_N_rate_from_holland_schepers()` directly.
- For crops with multiple phases (e.g. grano duro with fase nd), the MAS
  lookup uses the first matching row; pass `phase = ...` explicitly when
  ambiguity matters.

# Session info

```{r session}
sessionInfo()
```
