---
title: "Remote Sensing Diagnosis: multi-index VIs and Nitrogen Nutrition Index"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Remote Sensing Diagnosis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Overview

This vignette demonstrates the **remote-sensing diagnosis** pipeline added
in NFert 0.7.0:

1. Compute multiple vegetation indices (VIs) from a multi-band raster
   with `compute_vi()`.
2. Diagnose crop nitrogen status pixel-by-pixel with `compute_NNI()` /
   `diagnose_N_status()`, using the species-specific critical nitrogen
   dilution curve of Lemaire & Gastal (1997).
3. Feed the diagnosis into the variable-rate prescription pipeline
   (`variable_rate_N()`) so that the applied dose responds to the
   NNI diagnosis rather than to a raw vegetation index alone.

Why this matters: NDVI saturates at LAI > 3--4. In mid- to late-vegetative
stages (GS30+ for winter wheat; V8+ for maize) red-edge indices (NDRE,
CIred-edge) are far more sensitive to canopy N status. And all of those
indices are still just **proxies**: the scientifically correct diagnostic
variable is the NNI, which uses the dilution curve to convert biomass +
N% into a dimensionless deficiency / sufficiency index.

## 1. A synthetic multi-band stack (for the vignette)

In a real application the multi-band input comes from a Sentinel-2 L2A
scene or a UAV multispectral flight. Here we synthesise a small stack
so the vignette is reproducible offline.

```{r synthetic-stack}
library(NFert)
library(raster)

set.seed(1)
template <- raster(nrows = 40, ncols = 40,
                   xmn = 0, xmx = 100, ymn = 0, ymx = 100)

# Plausible canopy reflectances (fraction 0-1)
B03 <- setValues(template, runif(ncell(template), 0.05, 0.12))  # green
B04 <- setValues(template, runif(ncell(template), 0.04, 0.08))  # red
B05 <- setValues(template, runif(ncell(template), 0.10, 0.25))  # red-edge
B08 <- setValues(template, runif(ncell(template), 0.35, 0.55))  # NIR

s2 <- stack(B03, B04, B05, B08)
names(s2) <- c("B03", "B04", "B05", "B08")
```

## 2. Multi-index VI engine

`compute_vi()` computes any of NDVI, NDRE, GNDVI, CIred-edge, MCARI,
MSAVI2 from the stack, with configurable band mapping. Default mapping
follows Sentinel-2 L2A naming.

```{r compute-vi}
ndvi  <- compute_vi(s2, "NDVI")
ndre  <- compute_vi(s2, "NDRE")
gndvi <- compute_vi(s2, "GNDVI")
cire  <- compute_vi(s2, "CIred")

summary(getValues(ndvi))
summary(getValues(ndre))
```

```{r plot-vis, fig.width=7, fig.height=4}
par(mfrow = c(2, 2), mar = c(2, 2, 2, 3))
plot(ndvi,  main = "NDVI",  col = rev(terrain.colors(30)))
plot(ndre,  main = "NDRE",  col = rev(terrain.colors(30)))
plot(gndvi, main = "GNDVI", col = rev(terrain.colors(30)))
plot(cire,  main = "CIred", col = rev(terrain.colors(30)))
```

NDVI and GNDVI typically cluster in a narrow range for closed canopies
(saturation); NDRE and CIred-edge preserve variability at high biomass.
That is why the red-edge indices are the reference for N diagnosis in
GS30+ / V8+.

## 3. Critical-N curve

`critical_N_curve()` returns the (a, b, W_min) coefficients of the
species-specific critical nitrogen dilution curve. Italian synonyms
are accepted.

```{r critical-curve}
critical_N_curve("wheat")
critical_N_curve("mais")   # italian
```

The curve is Nc = a * W^(-b), where Nc is the critical N concentration
(% DM) and W the aboveground dry biomass (t DM / ha). Defaults come
from Justes 1994 (wheat), Plenet & Lemaire 2000 (maize), Sheehy 1998
(rice), Colnenne 1998 (rapeseed), Duru 1997 (grass), van Oosterom 2010
(sorghum), Debaeke 2012 (sunflower).

## 4. Scalar NNI diagnosis

```{r scalar-nni}
# Winter wheat at GS30: N% = 3.2, biomass = 2.5 t DM / ha
compute_NNI(N_content = 3.2, biomass = 2.5,
            crop = "wheat", is_percent = TRUE)
# -> ~0.94 : slightly deficient

# Maize at V8: N% = 2.8, biomass = 4 t DM / ha
compute_NNI(2.8, 4, crop = "maize", is_percent = TRUE)
# -> ~1.5  : luxury consumption
```

## 5. Raster-wise NNI diagnosis

In a realistic workflow biomass (W) and N concentration (N%) are both
raster layers, derived from remote sensing + empirical relationships.
Here we illustrate with two synthetic layers that emulate the output
of a biomass retrieval model + a leaf-N retrieval model.

```{r raster-nni, fig.width=7, fig.height=3}
# Synthetic biomass and N% layers
w_map <- setValues(template, runif(ncell(template), 1.5, 6.0))
n_map <- setValues(template, runif(ncell(template), 2.2, 3.9))

d <- diagnose_N_status(N_content = n_map, biomass = w_map,
                       crop = "wheat", is_percent = TRUE)

par(mfrow = c(1, 2), mar = c(3, 3, 2, 4))
plot(d$NNI,
     main = "NNI (continuous)",
     col  = hcl.colors(30, "RdYlGn"))
plot(d$class,
     main = "NNI class (1=deficient, 2=optimal, 3=excessive)",
     col  = c("#d73027", "#fee08b", "#1a9850"))
```

```{r nni-summary}
d$summary
d$thresholds
d$curve
```

## 6. Diagnosis-guided variable rate

The usual `variable_rate_N()` accepts any normalised VI. The
diagnosis-guided upgrade is: (a) use a red-edge index for the VRT, and
(b) zero out the prescription where the NNI diagnosis shows excessive
N status, so fertiliser is not applied where the crop is already
over-supplied.

```{r vrt-guided, fig.width=7, fig.height=3}
N_target <- 160   # e.g. from a preliminary N_balance()

vr <- variable_rate_N(ndre, n_dose = N_target,
                      method = "holland",
                      minN = 60, maxN = 200)

# Mask out pixels classified as "excessive" (class == 3)
rate_guided <- vr$rate_raster
rate_guided[d$class == 3] <- 0

par(mfrow = c(1, 2), mar = c(3, 3, 2, 4))
plot(vr$rate_raster,
     main = "VRT (NDRE + Holland)",
     col  = rev(heat.colors(30)))
plot(rate_guided,
     main = "VRT masked by NNI (class 3 -> 0)",
     col  = rev(heat.colors(30)))
```

```{r verify-guided}
cat("Full prescription  mean :",
    round(cellStats(vr$rate_raster, mean), 1), "kg/ha\n")
cat("NNI-guided        mean  :",
    round(cellStats(rate_guided, mean), 1), "kg/ha\n")
cat("N saved by NNI mask     :",
    round(cellStats(vr$rate_raster - rate_guided, mean), 1),
    "kg/ha avg, or",
    round(cellStats((vr$rate_raster - rate_guided), sum) *
          prod(res(rate_guided)) / 10000, 1),
    "kg total over the field\n")
```

## 7. Empirical NNI from a single VI raster (no GPR models)

The full `compute_NNI_from_S2()` pipeline requires the 8 pyeogpr
`.json` model files plus a 10-band Sentinel-2 L2A scene. When only a
single VI raster is available (UAV NDRE, satellite NDVI, etc.), the
helper `nni_from_vi_empirical()` delivers an NNI map directly through
a published linear regression:

```{r empirical-nni, eval = FALSE}
ndre <- raster::raster("ndre_field.tif")
out  <- nni_from_vi_empirical(
  ndre,
  index = "NDRE",       # or "NDVI", "CIred_edge"
  crop  = "wheat",      # wheat / maize / rice / barley
  # slope = 3.5, intercept = 0.30,  # override with local calibration
  nni_thresholds = c(0.90, 1.10))

raster::plot(out$NNI,   main = "NNI (empirical)")
raster::plot(out$zones, main = "Zones (1 deficient / 2 optimal / 3 excessive)")
```

Default slope and intercept are first-guess values from Cao 2013
(rice), Fitzgerald 2010 and Magney 2017 (wheat), Li 2014 (maize).
Local recalibration with 30-50 plot-level ground-truth samples is
strongly recommended.

## 8. From NNI to variable-rate strips

An NNI map (from either pipeline) can drive
`build_strip_prescription(variability = "nni", ...)` to produce
machine-width strips directly exportable to any tractor monitor:

```{r strip-from-nni, eval = FALSE}
ex    <- system.file("extdata/example_farm.geojson", package = "NFert")
field <- sf::st_read(ex, quiet = TRUE)[1, ]

rx <- build_strip_prescription(
  field         = field,
  machine_width = 24,
  nni_raster    = out$NNI,
  variability   = "nni",
  n_target      = 180, min_dose = 60, max_dose = 220)

export_prescription(rx, "rx_nni.shp", format = "shp")
```

## 9. Notes and limitations

* The VIs and NNI require a **closed canopy** (W >= W_min of the
  species-specific curve). NFert clamps W at W_min internally, but
  interpret outputs with care for early-vegetative imagery.
* `compute_vi()` does *not* perform cloud masking; clean the input
  stack beforehand (Sentinel-2 L2A ships a SCL layer that can be used
  to mask classes 3, 8, 9, 10 and 11).
* The diagnosis-guided VRT in section 6 is a didactic example. A
  production pipeline should combine the NNI classes with the MAS cap
  (`get_MAS()`) and the ZVN 170 kg N/ha check from `plan_distribution()`.

## References

Clarke, T.R. et al. (2001). Remote sensing of nitrogen status in wheat.
Proc. Beltwide Cotton Conf.

Duru, M. et al. (1997). Nitrogen nutrition status for forage grass
swards: application of the critical curve concept. Ann. Zootech. 46.

Justes, E. et al. (1994). Determination of a critical nitrogen
dilution curve for winter wheat crops. Ann. Bot. 74.

Lemaire, G. & Gastal, F. (1997). N uptake and distribution in plant
canopies. In: Diagnosis of the Nitrogen Status in Crops. Springer.

Li, F. et al. (2014). Improving estimation of summer maize nitrogen
status with red edge-based spectral vegetation indices. F. Crops Res. 157.

Plenet, D. & Lemaire, G. (2000). Relationships between dynamics of
nitrogen uptake and dry matter accumulation in maize crops. Plant Soil 216.

## Session info

```{r session}
sessionInfo()
```
